---
title: "Homework 6"
author: "Chloe Chah"
output: github_document
---

```{r setup, include = FALSE}
library(tidyverse)
library(viridis)
library(modelr)
library(mgcv)
library(p8105.datasets)
library(rnoaa)

knitr::opts_chunk$set(
	echo = TRUE,
	warning = FALSE,
	fig.width = 8, 
  fig.height = 6,
  out.width = "90%"
)
options(
  ggplot2.continuous.colour = "viridis",
  ggplot2.continuous.fill = "viridis"
)
scale_colour_discrete = scale_colour_viridis_d
scale_fill_discrete = scale_fill_viridis_d
theme_set(theme_minimal() + theme(legend.position = "bottom"))
```


## Problem 1

### Create a city_state variable and a binary variable indicating whether the homicide is solved. For this problem, limit your analysis those for whom victim_race is white or black. Be sure that victim_age is numeric.

```{r}

urlfile="https://raw.githubusercontent.com/washingtonpost/data-homicides/master/homicide-data.csv"

data <- read_csv(url(urlfile))

homicide_df = 
  data %>% 
  mutate(
    city_state = str_c(city, state, sep = ", "),
    victim_age = as.numeric(victim_age),
    resolution = case_when(
      disposition == "Closed without arrest" ~ 0,
      disposition == "Open/No arrest"        ~ 0,
      disposition == "Closed by arrest"      ~ 1)
  ) %>% 
  filter(
    victim_race %in% c("White", "Black"),
    city_state != "Tulsa, AL") %>% 
  select(city_state, resolution, victim_age, victim_race, victim_sex)
```


### For the city of Baltimore, MD, use the glm function to fit a logistic regression with resolved vs unresolved as the outcome and victim age, sex and race as predictors. Save the output of glm as an R object; apply the broom::tidy to this object; and obtain the estimate and confidence interval of the adjusted odds ratio for solving homicides comparing non-white victims to white victims keeping all other variables fixed.

```{r}
baltimore_df =
  homicide_df %>% 
  filter(city_state == "Baltimore, MD")
glm(resolution ~ victim_age + victim_race + victim_sex, 
    data = baltimore_df,
    family = binomial()) %>% 
  broom::tidy() %>% 
  mutate(
    OR = exp(estimate),
    CI_lower = exp(estimate - 1.96 * std.error),
    CI_upper = exp(estimate + 1.96 * std.error)
  ) %>% 
  select(term, OR, starts_with("CI")) %>% 
  knitr::kable(digits = 3)
```


### Now run glm for each of the cities in your dataset, and extract the adjusted odds ratio (and CI) for solving homicides comparing Black victims to white victims. Do this within a “tidy” pipeline, making use of purrr::map, list columns, and unnest as necessary to create a dataframe with estimated ORs and CIs for each city.

```{r}
models_results_df = 
  homicide_df %>% 
  nest(data = -city_state) %>% 
  mutate(
    models = 
      map(.x = data, ~glm(resolution ~ victim_age + victim_race + victim_sex, data = .x, family = binomial())),
    results = map(models, broom::tidy)
  ) %>% 
  select(city_state, results) %>% 
  unnest(results) %>% 
  mutate(
    OR = exp(estimate),
    CI_lower = exp(estimate - 1.96 * std.error),
    CI_upper = exp(estimate + 1.96 * std.error)
  ) %>% 
  select(city_state, term, OR, starts_with("CI")) 
```

### Create a plot that shows the estimated ORs and CIs for each city. Organize cities according to estimated OR, and comment on the plot.

```{r}
models_results_df %>% 
  filter(term == "victim_sexMale") %>% 
  mutate(city_state = fct_reorder(city_state, OR)) %>% 
  ggplot(aes(x = city_state, y = OR)) + 
  geom_point() + 
  geom_errorbar(aes(ymin = CI_lower, ymax = CI_upper)) + 
  theme(axis.text.x = element_text(angle = 90, hjust = 1))
```



## Problem 2

### Load and clean the data for regression analysis (i.e. convert numeric to factor where appropriate, check for missing data, etc.).

```{r}
baby_df = 
  read_csv("./Data/birthweight.csv", na = c("", "NA", "Unknown")) %>%
  mutate(
    frace = as.factor(frace), 
    mrace = as.factor(mrace),
    babysex = as.factor(babysex),
    malform = as.factor(malform)
  )

any(is.na(baby_df))

head(baby_df)

```

There appears to be no missing data in this dataset. I've converted data types that could be a combo box (e.g., sex, father's race, mother's race, and presence of malformations that could affect weight) from numeric to factors. 

### Propose a regression model for birthweight. This model may be based on a hypothesized structure for the factors that underly birthweight, on a data-driven model-building process, or a combination of the two.

```{r}
model_fit = lm(bwt ~ babysex + gaweeks + parity + ppwt + wtgain + momage, data = baby_df)
```

I included maternal factors (number of live births prior to this pregnancy,mother’s pre-pregnancy weight, mother's weight gain during pregnancy, and mother's age at delivery) because based on recent publications, it appears that health conditions and characteristics of mothers before and during pregnancy may directly affect the baby's birthweight. 

### Show a plot of model residuals against fitted values – use add_predictions and add_residuals in making this plot.

```{r}
baby_df %>% 
  add_predictions(model_fit) %>%
  add_residuals(model_fit) %>% 
  ggplot(aes(x = pred, y = resid)) + 
  geom_point() + 
  labs(title = "Fitted values vs. Residuals", 
      x = "Fitted values", 
      y = "Residuals")
```

Based on the residuals plot, the residuals appear to be relatively evenly distributed. The variance of the residuals appear to be the same for the most part. That is, the homoscedasticity assumption seems to be met and the values are concentrated in the center and distributed in a rectangular pattern (randomly scattered about a horizontal line). There doesn't seem to be any systematic pattern or clustering that is considered a violation of the homoscedasticity assumption. 

### Compare your model to one using length at birth and gestational age as predictors (main effects only) and to one using head circumference, length, sex, and all interactions (including the three-way interaction) between these. 

```{r error = TRUE}

cv_df = 
  crossv_mc(baby_df, 100) %>%
  mutate(
    model_fit = map(train, ~lm(bwt ~ babysex + gaweeks + parity + ppwt + wtgain + momage, data = baby_df)), 
    main_fit = map(train, ~lm(bwt ~blength + gaweeks, data = baby_df)), 
    interaction_fit = map(train, ~lm(bwt ~ bhead + blength + babysex + bhead*blength + bhead*babysex + blength*babysex + bhead*blength*babysex, data = baby_df))) %>%
  mutate(
    rmse_model = map2_dbl(model_fit, test, ~rmse(model = .x, data = .y)), 
    rmse_main = map2_dbl(main_fit, test, ~rmse(model = .x, data = .y)), 
    rmse_interaction = map2_dbl(interaction_fit, test, ~rmse(model = .x, data = .y))
  )

cv_df %>% 
  select(starts_with("rmse")) %>% 
  pivot_longer(
    everything(),
    names_to = "model", 
    values_to = "rmse",
    names_prefix = "rmse_") %>% 
  mutate(model = fct_inorder(model)) %>% 
  ggplot(aes(x = model, y = rmse)) + 
  geom_violin() + 
  labs(
    x = "Model", 
    y = "RMSE", 
    title = "Distribution of RMSE values for each candidate model"
  )

```

Based on these results, the model that I proposed has the highest RMSE values compared to the main effects and interaction models. This could be due to the fact that my model mainly included parental factors. If I included more factors directly related to baby's birthweight such as baby's head circumference at birth, my model might have performed better. In fact, it appears that the interaction model is the best fit for this dataset because lower RMSE values indicate better fit.  

## Problem 3

### Load the data
```{r error = TRUE}

weather_df = 
  rnoaa::meteo_pull_monitors(
    c("USW00094728"),
    var = c("PRCP", "TMIN", "TMAX"), 
    date_min = "2017-01-01",
    date_max = "2017-12-31") %>%
  mutate(
    name = recode(id, USW00094728 = "CentralPark_NY"),
    tmin = tmin / 10,
    tmax = tmax / 10) %>%
  select(name, id, everything())

head(weather_df)

```

### Use 5000 bootstrap samples and, for each bootstrap sample, produce estimates of r^2 and log(β^0∗β^1)

```{r}

bootstrap_df = 
  weather_df %>%
  modelr::bootstrap(n = 5000) %>% 
  mutate(
    models = map(strap, ~lm(tmax ~ tmin, data = .x) ),
    results_r = map(models, broom::glance), 
    results_b = map(models, broom::tidy)) %>%
  janitor::clean_names()

r_squared = 
  bootstrap_df %>%
  select(-strap, -models) %>% 
  unnest(results_r) %>% 
  janitor::clean_names() %>%
  select(id, r_squared)

log_beta = 
  bootstrap_df %>%
  select(-strap, -models) %>% 
  unnest(results_b) %>% 
  select(id, term, estimate) %>%
  pivot_wider(
    names_from = term, 
    values_from = estimate)%>%
  janitor::clean_names() %>%
  mutate(
    log_b0_b1 = log(tmin*intercept)
  )

```

### Plot the distribution of r squared values 

```{r}

r_plot = 
  r_squared %>%
  ggplot(aes(x = r_squared)) +
  geom_density() + 
  labs(
    title = "Distribution of r-squared", 
    x = "R-squared", y = "Count"
  )

r_plot

```

Based on the distribution plot, it appears that the R-squared values gathered from bootstrapping seem to be approximately normally distributed or symmetric around a mean of 0.91. 

### Plot the distribution of log(b0*b1)
```{r}

b_plot = 
  log_beta %>%
  ggplot(aes(x = log_b0_b1)) +
  geom_density() + 
  labs(
    title = "Distribution of log(b0*b1)", 
    x = "log_b0_b1", y = "Count"
  )

b_plot

```

Based on the distribution plot, it appears that the log(β^0∗β^1) values gathered from bootstrapping seem to be approximately normally distributed around a mean of 2.01. Compared to the distribution plot for r-squared values, it does seem a little bit more skewed to the left. 


### Using the 5000 bootstrap estimates, identify the 2.5% and 97.5% quantiles to provide a 95% confidence interval for r^2 and log(β^0∗β^1)

```{r}

log_beta %>%
  summarize(
    ci_lower = quantile(log_b0_b1, 0.025), 
    ci_upper = quantile(log_b0_b1, 0.975)
  ) %>%
  knitr::kable()

r_squared %>%
  summarize(
    ci_lower = quantile(r_squared, 0.025), 
    ci_upper = quantile(r_squared, 0.975)
  ) %>%
  knitr::kable()

```

The 95% confidence intervals for the log(β^0∗β^1) values is (1.964469, 2.057902). 
The 95% confidence intervals for the r-squared values is (0.8938043, 0.9270294)


